Stack Safety is a Security Property

### 2021-04-15 tex/main.tex
Nits

We propose formal criteria for stack safety, using the concepts of
language-based security: a combination of an integrity property (``each
caller's stack frame is held invariant by the callee'') and a confidentiality
property (``the callee's behavior is insensitive to differences in the
caller's state'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We argue that these criteria match common intuitions by (1) showing that
they are violated by a variety of intuitively unsafe examples and (2)
using property-based random testing to check that appropriate variants of
the criteria are actually enforced by the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
broken variants).  Specifically, Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a basic ``stepwise'' versions of our confidentiality
and integrity properties. Their more performant ``lazy'' micro-policy
does not, but corresponds to weaker, more extensional ``observational'' variants
of the
properties.  While we do not rule out other enforcement mechanisms, our
definitions are strongly informed by micropolicy systems, especially
our attacker model, which is stronger than other enforcement mechanisms
can enable.

@@

In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This is novel in that the
confidentiality property applies to many nested subtraces within a single execution,
as opposed to traces of entire executions.
{\em Integrity} says that a caller's local data is protected from modification by its callee,
motivated by the reasoning principle that local data persists across a call. Its
{\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. Once again
{\em stepwise} describes the granularity of enforcement: at every step of a call,
changes in the state are independent of the caller's data.

Together, these properties capture the rough intuition of LSE, but are somewhat stronger.
First, we model returns as returns to a valid return target; an invalid return
is treated as a mere jump within the callee, and does not grant access to the caller's
data. So, even without WBCF, our stack safety properties rule out many control flow attacks.
The attacker in such an instance is permitted to ``return'' so long as it avoids
violating the integrity and confidentiality of the caller. Second, confidentiality
contains temporal protections that LSE does not necessarily address. The intuition
for LSE is that it protects the caller's stack frame, not necessarily memory that was
previously allocated but now outside of the caller's frame. (Cheri-capability-based
implementations of LSE {\em do} protect previously allocated data, typically by clearing
it, but because it is a potential vector for leaking capabilities into the caller's frame,
not as a goal of the property.)

To illustrate the applicability of these properties, we perform randomized testing to
validate that an existing enforcement mechanism does enforce them.
\citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}, but offer no formal properties.
We test against their {\em Depth Isolation} micro-policy, in which memory cells within each stack
frame are tagged with the identity of the function activation
that owns the frame, and access to those locations is then gated
according to the activation that is currently executing. This is simple and corresponds
to intuitive notions of stack safety, and we argue that by describing its behavior
we approach a reasonable stack-safety property that other micro-policies can be
validated against.

Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If we can catch known bugs, we gain confidence that we can catch unknown ones,
strengthening the validation of the enforcement mechanism.

@@

A second challenge is that enforcement mechanisms may weaken their protections
in the name of efficiency, while arguing that they still capture stack safety in a meaningful sense.
\citet{DBLP:conf/sp/RoesslerD18} explore alternative micro-policies, with different
ways of identifying functions and deciding exactly which accesses are allowed.
These vary in their efficiency and in the spatial and temporal precision of their
security guarantees.
The Depth Isolation policy prevents a callee from writing outside its frame
at all, which is simple and intuitively correct, but it must manually clear the tags
on the frame on each return, which is very expensive. Conversely, their
\emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. This violates our
straightforward properties, but ntuitively,
this form of laziness is justified if any dangerous writes will
ultimately be detected before they can affect program behavior.

What does ``affecting program behavior'' really mean? This micro-policy
calls for a second characterization of stack safety in terms of the externally
observable behavior of a stack-safe system. Such a characterization will have the
additional benefit of being more {\em extensional}, concerned primarily with the
external behavior of the system. In contract, our stepwise properties are {\em intensional},
focusing on internal details that may not actually matter to overall security.

To formalize this intuition, we develop an \emph{observational integrity} property which
characterizes enforcement mechanisms that detect violations between the point when
they occur and the point when they make an observable impact. Intuitively this property
can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
integrity implies observational integrity. \sna{TODO: check this.}

We additionally sketch an associated {\em observational confidentiality} property
makes confidentiality extensional in the same way, in order to form an
{\em observational stack safety} that is highly extensional.
We therefore argue that, despite its somewhat greater complexity,
it may be a better minimal specification for
stack safety enforcement mechanisms in general.

%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
As it turns out, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy enforces observational integrity, but not confidentiality,
whether stepwise or observational.
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality occurs in corner cases where callees
can leak data between one another.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
confidentiality, albeit at some performance cost.

Our initial description of these properties assumes a single simple stack with no sharing
between callers and callees: all parameters and return values are passed in registers.  Later,
we show how to refine it to allow passing of scalar stack data, granting a
callee access to the caller's data. We do not model more sophisticated sharing,
such as taking the address of a variable and treating it like a pointer.

@@

\section{Setup}
\label{sec:setup}

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it.  We assume a register
machine with a program counter (\(\PCname\)); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including where code lives and which addresses are within
the bounds of the stack. We do not model a heap; programs might choose to use memory
that is neither code nor stack to store global variables, as a heap, etc., but our
properties do not protect it or assume any particular behavior. We also do not
model run-time code generation.

Our stack-safety criteria depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, and we determine when a return has occurred by tracking
control flow and observing when we reach a {\em return target}. For a given state annotated
as a call, in a standard calling convention, another state fulfills its return target if
it has the same stack pointer and the program counter is at the instruction following
the call, but we abstract over different conventions.
%, returns, and function entry points\BCP{I
%  still want to talk about whether returns and entry points need to be
%  annotated.  To me, these annotations do not make intuitive sense: it is
%  the caller's responsibility to know whether the address that it is calling
%is really an entry point, and we can trust it to check this.  Similarly, it
%is not the callee's role to say when it is returning; it is the caller's
%prerogative to decide when it has been returned to!  I.e., from the caller's
%point of view, it has been returned to when control resumes at the
%instruction following the call with the SP reset to where it was at the
%point of the call. (This is the point where it wants to be able to access
%its own local state again and expected to find it unmolested.)}.
Although call annotations would typically come from a compiler together with
the machine code, we do not assume anything about their provenance.

@@

Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

The separation is smaller in protection mechanisms that interact more directly
with application code. Some mechanisms might exhibit failstop behavior, but have
no protected policy state; others might never even failstop. Then the policy-enhanced
machine is simply identical to the original. This is the case for software-based
techniques.

\section{Threat Model}
\label{sec:threat}

@@

\BCP{I believe we could delete this section---and never mention WBCF at all
  (except to say that we leave it out, compared to STkTokens)---if we
  eliminated return (and entry point) annotations.}
\sna{I'm not sure we can escape questions about annotations, even if we have
  just calls, but I tried to abbreviate it.}

Since we rely heavily on call annotations, we must ask: what if the annotations
are wrong? First, suppose a call occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call, only an instruction sequence resembling one, and
therefore no associated change in permissions. Then the caller is guaranteed no
protections.

Alternatively, an annotated instruction might not obey the calling convention.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish, termed {\em blessed sequences}.
When applying our properties to a micro-policy setting, we will statically rule out some
scenarios with a well-formedness criterion on initial states: all annotated instructions will
be situated at the appropriate instruction of a blessed sequence, tagged according to the policy.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.

@@

%This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
%We believe that it is flexible enough to capture code-altering approaches, such as code
%rewriting or capability-based techniques, though perhaps with less elegance.

@@

\begin{subfigure}{.2\textwidth}
{\small
\begin{verbatim}
int main() {
  int x = 42;
  return f()+x;
}

int f() {
  int y;
  *(&y-2) = 0;
  return 5;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
{\small
\begin{verbatim}
1 main: add $1,%sp      ; allocate frame
2       mov $42,-1(%sp) ; initialize local
3       call f
4       add -1(%sp),%ra ; set return value
5       sub $1,%sp      ; deallocate frame
6       ret

10 f:   add $1,%sp      ; allocate frame
11      mov $0,-3(%sp)  ; violation!
12      mov $5, %ra     ; set return value
13      sub $1,%sp      ; deallocate frame
14      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.25\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 2} &
\memoryaddrs{8em}
\memory{3}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 3} &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 12} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}

@@

This section will elucidate the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will argue that:

@@

We show our examples both in C syntax and in corresponding assembly code for a
simple two-address machine. The address of each instruction is shown to its left.
The assembly code reflects a simple compilation model in which  
local variables are stored in the stack frame.
Unless otherwise stated, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \$SP} points to the first unused word above the top of the stack.
The {\tt call} and {\tt ret} instructions push and pop the return address from the stack,
implicitly adjusting the stack pointer.
At the right of each example, we show the layout of stack memory just before
the specified instruction is executed; each box represents a memory word and
addresses increase to the right.

Unless otherwise noted, colored boxes indicate an assignment of the address to a
{\em domain} -- a label reflecting a particular set of access rights. Domains
are \(\unsealed\) (in the stack and accessible to writes), \(\sealed{\depth}\)
(in the stack and reserved for the caller at depth \(\depth\)), and \(\outside\)
(outside the stack entirely.) For our purposes, \(\outside\) will apply to global
variables. Domains are part of the {\em context} of a state.
We will introduce the remaining context as needed.

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} with {\tt main}'s depth,
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after instruction 8, {\tt f} has violated integrity by writing to {\tt x},
so the final return value produced by {\tt main} is wrong.

The black rectangle represents the return address, which is saved by the call
instruction itself, so is not sealed.

\begin{figure}
\begin{subfigure}{.2\textwidth}
{\small
\begin{verbatim}
int main() {
  int x = 42;
  return f()+x;
}

int f() {
  int y,z;
  y = 5;
  z = *(&y-2);
  return y+z;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
{\small
\begin{verbatim}
1 main: add $1,%sp      ; allocate frame
2       mov $42,-1(%sp) ; initialize local
3       call f
4       add -1(%sp),%ra ; set return value
5       sub $1,%sp      ; deallocate frame
6       ret

>>>>

We propose formal criteria for stack safety, using the concepts of
language-based security: a combination of an integrity property (``each
caller's stack frame is held invariant by the callee'') and a confidentiality
property (``the callee's behavior is insensitive to differences in the
caller's state'').
% These properties are sensible even in the presence of
% abnormal control flow.\apt{Is this important enough to mention here?}
% BCP: I think not in the abstract, but maybe in the intro.
% We further apply\apt{what does this mean? we didn't
%  invent it} a straightforward notion of
% ``well-bracketed control flow'' that restricts control flow to the
% typical pattern of matched calls and returns.
%
We argue that these criteria match common intuitions by (1) showing that
they are violated by a variety of intuitively unsafe examples and (2)
using property-based random testing to check that appropriate variants of
the criteria are actually enforced by the stack-safety ``micro-policies''
proposed by~\citet{DBLP:conf/sp/RoesslerD18} (and {\em not} enforced by
broken variants).  Specifically, Roessler
and Dehon's ``eager'' micro-policy, which catches violations as early as
possible, enforces a basic ``stepwise'' versions of our confidentiality
and integrity properties. Their more performant ``lazy'' micro-policy
does not, but corresponds to weaker, more extensional ``observational'' variants
of the
properties.  While we do not rule out other enforcement mechanisms, our
definitions are strongly informed by micropolicy systems, especially
our attacker model. \BCP{I am still not clear in what way, exactly.}

@@

In applying these security concepts, we treat each caller as needing protection
from its callee, even as that callee may need protection from its callee in turn.
This differs from common practice in language-based security in that the
confidentiality property applies to many nested subtraces within a single execution.\sna{How novel is this nesting?}\bcp{I believe completely novel?}
{\em Integrity} says that a caller's local data is protected from modification by its callee,
motivated by the reasoning principle that local data persists across a call. Its
{\em stepwise} interpretation formalizes this very naturally: a caller's local data
is never modified during a call. {\em Confidentiality} says that a caller's data is protected
from being leaked into its callee, and also that the callee cannot see uninitialized data.
As a reasoning principle, confidentiality guarantees that the callee's
behavior is determined only by a predictable interface that does not include the caller's
local state or the uninitialized portion of the stack. This is formalized as a
{\em noninterference}-style property that, rather than comparing pairs of whole program traces,
compares a pair of subtraces for each call, nested within other calls. Once again
{\em stepwise} describes the granularity of enforcement: at every step of a call,
changes in the state are independent of the caller's data.

Together, these properties capture the rough intuition of LSE, but are somewhat stronger.
First, we model returns as returns to a valid return target; an invalid return
is treated as a mere jump within the callee, and does not grant access to the caller's
data. So, even without WBCF, our stack safety properties rule out many control flow attacks.
The attacker in such an instance is permitted to ``return'' so long as it avoids
violating the integrity and confidentiality of the caller. Second, confidentiality
contains temporal protections that LSE does not implicitly address.
\apt{Maybe want to say a little more about this here, since it is a novelty wrt/ existing work?}
\bcp{+1}\sna{TODO: Say a little more, but make sure it's accurate first.}

@@

A second challenge is that
enforcement mechanisms may weaken their protections in the name of efficiency,
while arguing that they still capture stack safety in a meaningful sense.
For example, \citet{DBLP:conf/sp/RoesslerD18} describe how to protect the stack
using tag-based \emph{micro-policies}~\citep{pump_oakland2015}.
Memory cells within each stack frame are tagged with the
identity of the function activation
that owns the frame\BCP{Really??}, and access to those locations is then gated
according to the activation that is currently executing. Roessler and DeHon
explore alternative versions of this basic idea, with different ways of identifying
functions and deciding exactly which accesses are allowed.
These vary in their efficiency and in the spatial and temporal precision of their
security guarantees.
The \emph{Depth Isolation} policy prevents a callee from writing outside its frame
at all, which is simple and intuitively correct, but it must manually clear the tags
on the frame on each return, which is very expensive. Conversely, the
\emph{Lazy Tagging and Clearing} policy improves
performance by assigning locations to the active function when they are written, which
permits  ``illicit'' writes to inaccessible stack frames so long as the
written locations are not subsequently read by their ``real'' owners. Intuitively,
this form of laziness is justified if any dangerous writes will
ultimately be detected before they can affect program behavior.

But what does ``affecting program behavior'' really mean? This micro-policy
calls for a second characterization of stack safety in terms of the externally
observable behavior of a stack-safe system. Such a characterization will have the
additional benefit of being more {\em extensional}, concerned primarily with the
external behavior of the system. In contract, our stepwise properties are {\em intensional},
focusing on internal details that may not actually matter to overall security.

To formalize this intuition, we develop an \emph{observational integrity} property which
characterizes enforcement mechanisms that detect violations between the point when
they occur and the point when they make an observable impact. Intuitively this property
can be phrased as ``no visible reads after dangerous writes.''  We prove (in Coq) that stepwise
integrity implies observational integrity.
\sna{There is a bit of massaging old proofs to do here, and I don't recall if we
  established that this was important.\apt{medium importance. it is obvious that
    observational integerity allows behaviors that stepwise does not, but
    maybe not obvious that the converse is not true too.}}
We additionally sketch an associated {\em observational confidentiality} property
makes confidentiality extensional in the same way, although we do not know of a real
enforcement mechanism that weakens confidentiality in this way.
\apt{why bother mentioning it then?}
\sna{If a relatively extensional characterization is a worthwhile contribution
  in its own right, then it's worth mentioning that we have one. If not, then indeed,
  we shouldn't mention it -- or get too excited about extensionality in general, because
  lockstep isn't extensional.}
Together, these form an {\em observational stack safety} that is highly extensional.
We therefore argue that, despite its somewhat greater complexity,
it may be a better minimal specification for
stack safety enforcement mechanisms in general.

\BCP{This paragraph doesn't apply to the lazy properties, so maybe it should
be moved before they are introduced?}%
To illustrate the applicability of these properties, we perform randomized testing to
validate that \citeauthor{DBLP:conf/sp/RoesslerD18}'s Depth Isolation micro-policy enforces
them. Randomized testing is a relatively low-cost way to validate a potential enforcement
mechanism; it finds flaws in the enforcement fast enough to be integrated into the workflow
of enforcement development \citep{TestingNI:ICFP}. To demonstrate that the testing adequately
covers potentially erroneous programs, we test intentionally buggy policies alongside
the real one. If the testing can differentiate correct implementations from buggy ones,
it is an argument in favor of their correctness. \bcp{Hmm--don't understand
  that sentence}
%\apt{more to say?}\bcp{Yes---partly to underscore our claim that this is a
%  significant contribution, and partly to give the reader some context for
%  why this is part of our story.  E.g., ``One nice thing we can do with our
%  spec is to use it to evaluate the claims of existing or future proposals
%  for enforcement mechanisms.  One relatively inexpensive way to carry out
%  such an evaluation is to see whether property-based testing against this
%  spec can distinguish correct implementations of the mechanism from buggy
%  ones...''}
Of course, as our testing reveals, \citeauthor{DBLP:conf/sp/RoesslerD18}'s Lazy Tagging and
Clearing micro-policy does not enforce stepwise integrity.
Unfortunately the existing Lazy Tagging and Clearing
micro-policy does not quite enforce observable confidentiality either. %, as we show in
%testing~\apt{really??},
But formalizing the ideal property allows us to see how it falls short -- a specific
violation of the temporal aspect of confidentiality.
We propose  a variant of Lazy Tagging and Clearing that we believe \emph{does} enforce
observational integrity, at some performance cost.\apt{but we're not testing that, right?}
\rb{Not sure about the property part right now, the machine side of this is
  pretty much done}
\sna{It's aspirational so far, but there's no fundamental reason we can't test it. We may
  need a caveat about how it's harder to test?}\apt{As per email discussion,
  maybe we could have a property variant that \emph{does} match depth-based Lazy Tagging and
  Clearing, by loosening the temporal safety properties?}
\sna{I'm pretty sure for the main policy it's just lazy integrity, no confidentiality.
  There was a variant Nick and Andre gave that would restrict the leaks to only be between
  instances of the same function, so that's the same property under a weaker attacker model,
  where we don't protect a function from itself.}

Our initial description of these properties assumes a single simple stack with no sharing
between callers and callees: all parameters and return values are passed in registers.  Later,
we show how to refine it to allow passing of scalar stack data, granting a
callee access to the caller's data. We do not model more sophisticated sharing,
such as taking the address of a variable and treating it like a pointer. Doing so without
protecting the variable is not interesting, and protecting would be better suited to
a different kind of enforcement, such as a memory safety micro-policy or Cheri-style capability.
\bcp{This is all a bit mysterious} \sna{Any less mysterious now?}

\rb{So no pointer passing? Pass by reference is now discussed at the end of
  Section 3, however it looks like things get tricky rather quickly.}
\sna{The trick is that if a programmer is passing pointers, we can't presume their
  intent -- how long should we permit memory to be shared? But if the compiler is passing
  pointers, as in pass by reference, we have a clear sense of where sharing should stop.
  \apt{N.B.Doing pass-by-reference properly requires that the passed variable is accessible
    by all descendents, not just immediate callee, which is doable with depth tags but not
    obviously with per-activation tags }}

@@

\section{Set Up and Threat Model}
\label{sec:threat}

\newcommand*{\rsp}{\textsc{sp}}

Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it.  We assume a register
machine with a program counter (\(\PCname\)); other
details of the machine's ISA are unimportant. We assume that we know ahead of time
the layout of memory, including where code lives and which addresses are within
the bounds of the stack.\bcp{This implies no run-time code generation,
  right?  This should probably be called out.}

\BCP{What should we say about the heap?}

Our stack-safety criteria depend
critically on the notions of ``call'' and ``return,'' which are not fully explicit
in machine code---the same instruction opcode may be used in some places as part of
a call or return sequence and in other places for other purposes.
%
So we assume we are given a machine language program {\em annotated} with the locations of
instructions representing calls, returns, and function entry points\BCP{I
  still want to talk about whether returns and entry points need to be
  annotated.  To me, these annotations do not make intuitive sense: it is
  the caller's responsibility to know whether the address that it is calling
is really an entry point, and we can trust it to check this.  Similarly, it
is not the callee's role to say when it is returning; it is the caller's
prerogative to decide when it has been returned to!  I.e., from the caller's
point of view, it has been returned to when control resumes at the
instruction following the call with the SP reset to where it was at the
point of the call. (This is the point where it wants to be able to access
its own local state again and expected to find it unmolested.)}. Although these
annotations would typically come from a compiler together with the machine code,
we do not assume anything about their provenance.

\bcp{Can this be postponed to the formal sections?}We further parameterize over the structure of the stack, particularly the means by
which a caller identifies the data that should be protected during a call. We term
this the {\em sealing convention}, which maps a machine state to a set of addresses.
If a state's \(\PCname\) is annotated as a call, the sealing convention tells us
which addresses must be protected until the return.

@@

\sna{This is awkward.}
This division of policy state from machine state is helpful when reasoning about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this,
with tags and extra policy state separated from and untouchable by application code.
Another significant example is the shadow stack, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

The separation is less useful\bcp{but still sensible, right} in protection mechanisms that interact more directly
with application code, such as Cheri-style capabilities\citep{??}\bcp{This
  section would read better if we could cleanly separate the general
  definitions from examples}. Here
the policy would not have auxiliary state, but would still refine the machine step
function to create a distinct failstop condition. Other techniques such as
software fault isolation~\citep{??} might never actually failstop, but fulfill
the properties by rewriting code so that traces can never violate them.

\paragraph*{Threat Model}

@@

\BCP{I believe we could delete this section---and never mention WBCF at all
  (except to say that we leave it out, compared to STkTokens)---if we
  eliminated return (and entry point) annotations.}

Since we rely heavily on code annotations, we must ask: what if the annotations
are wrong? First, suppose a call or return occurs that is not annotated.
In this case, the answer is simple: as far as the properties are concerned, there has
been no call or return, only an instruction sequence resembling such an operation, and
therefore no associated change in permissions. This is particularly interesting in the
case of a false return, where control may appear to have returned to the caller, but
the enforcement mechanism must still protect the caller's memory.

Alternatively, an annotated instruction might not really be a call or return.
In the context of \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies, well-behaved calls
and returns are implemented by fixed sequences of instructions which the tag policy
will force to be executed from start to finish. The annotated call or return instruction should always
be the final instruction of the appropriate sequence. What if an annotated instruction exists outside
such a sequence? We will statically rule out some scenarios with a well-formedness criterion
on initial states: annotated instructions will initially be tagged apropriately, as the
final instruction in the call or return sequence, and all blessed\bcp{undefined} sequences will end with
annotated instructions. This is all we need for the policy to
protect itself: it can then failstop if it would execute the blessed sequence out of order.
So, even in our strong attacker model, we need very little in the way of restrictions on
the code when considering micropolicy-based enforcement.  \BCP{Is this
  discussion still needed?  I had hoped that switching to our ``new''
  definition on subtraces would eliminate such complexities!}

\BCP{Not sure what this is about.}In general, regardless of enforcement mechanism, we take this approach:
a given enforcement mechanism may place a static well-formedness criterion
on the code it applies to. Weaker enforcement mechanisms might also need dynamic
reasonableness conditions to prevent the defender from sabotaging enforcement, as
discussed in section \ref{sec:cheri}.  \BCP{Weak.}

@@

This enforcement model directly describes micro-policies~\citep{pump_oakland2015}.
We believe that it is flexible enough to capture code-altering approaches, such as code
rewriting or capability-based techniques, though perhaps with less elegance.

@@

\begin{subfigure}{.4\textwidth}
\begin{verbatim}
int main() {
0:     int x, y;
1:     y = 42;
2[C]:  x = f();
3[R]:  return x+y;
}

int f() {
10:     *($SP-2) = 0;
11[R]:  return 5;
}
\end{verbatim}
\end{subfigure}
\begin{subfigure}{.4\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 1} &
\memoryaddrs{12em}
\memory{4}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 2} &
\memoryaddrs{12em}
\memory{4}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 10} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{2}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}

@@

This section will elucidate the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will insert our model between the enforcement and
the examples\bcp{Didn't understand that}, arguing that:

@@

Our examples are written in C-like syntax, labeled with line numbers for reference,
but to reflect the power of the underlying assembly code, we allow code
to read the stack pointer ({\tt \$SP}) directly and use
it to form pointers.
Call and return annotations are marked {\tt[C]} and {\tt[R]}
respectively. Local variables are assumed to be stored in the stack frame,
in declaration order. \bcp{This sort of thing implies that readers will have
to mentally compile the source programs we're showing them.}Unless otherwise stated, arguments and return values are passed in registers.
The machine is word-addressed. The stack grows upward in
memory and {\tt \$SP} points to the first unused word above the top of the stack.
At the right of each example, we show the layout of stack memory just before
the specified line number is executed; each box represents a memory word and
addresses increase to the right.  \BCP{I am worried about the C-like syntax:
I think the mental complexity and imprecision that it introduces may not be
worth the concision that we get from it.  Couldn't we get most of the
concision by just assuming an ISA with CISC-style single-instruction Call
and Return?}

%The boundary between the frames of a caller and callee is indicated by the $\rsp$ register
%at the time of the call. The stack can be used to pass arguments or return results
%in the usual way, by placing them at the top of the caller's frame. This will be modeled
%as a form of sharing in section \sna{cite the sharing section}.
%Callers and callees can also communicate via registers if they wish; indeed, our
%definitions say nothing at all about registers (except $\PCname$ and $\rsp$).

%\iftrue\bcp{This implicitly amounts to choosing a caller-saves register
%  convention, though, doesn't it?\apt{I think it is neutral; we could add separate caller register protection if we wanted.}}\fi

Figure~\ref{fig:int1} shows a simple integrity violation. It consists of a {\tt main}
function that performs a call to a function {\tt f}, which overwrites {\tt main}'s
variable {\tt x}. In the initial memory layout the entire stack is {\em unsealed},
meaning that it is eligible to be written to. At the call from {\tt main} to
{\tt f}, {\tt main}'s locals are {\em sealed} and labeled with the depth of
their owner\bcp{are we still using depths?},
in this case 0. Sealing is the fundamental integrity contract between the caller and callee:
{\tt main} expects that {\tt x} and {\tt y} will be unchanged when it gets control back.
This gives us our simple, intuitive statement of {\em stepwise integrity}:
if a component is sealed before a step, it will be unchanged afterward. Stepwise
refers to the fact that the property must hold on every step. %Later we will
%loosen this condition in the interest of extensionality -- the caller only
%cares about changes it can see after the return, not changes during execution.
In this case, after line 10, {\tt f} has violated integrity by writing to {\tt x}.

The black rectangle represents the return address. In a typical RISC calling convention
this will have been saved into a register by the call itself, and {\tt f} stores
it in the stack during its preamble, so it is not itself sealed.

\begin{figure}
\begin{subfigure}{.33\textwidth}
\begin{verbatim}
int main() {
0:     int x;
1:     x = 42;
2[C]:  x = f();
3[R]:  return x+y;
}
